### 2023-06-22 python_excel_ite.tex
working on intro

Despite numerous calls for and predictions of the eradication of spreadsheets from the business world, Google "spreadsheets are evil"

Excel not a database and shouldn't be used as one. There are better data visualizaiton and exploration tools. There are better dashboarding tools. Spreadsheet errors are a huge problem. Version control and documentation issues. Data size limitations. External link problems. But, spreadsheets are still very well suited to what if type modeling due to their flexibility, numerous built in functions, integrated visualization capabilities and ubiquity

Advent of programmatic analytics using tools like R and Python have changed the teaching landscape. Motivated by reproducibility, open source, free, and ease of integration with web and cloud databases.

Created PCDA course in response. Challenges with finding good textbooks. Developed much of my own curricular materials. Taught using combo of R Markdown documents and Jupyter notebooks. Literate computing \cite{Jupyter_proposal}. Educational narratives that support iterative exploration and development. The ability to weave standard language with formal programming languages is critical to creating the desired educational artifact. Jupyter used by many universities as a key component of their educational technology stack.

PCDA course focused mostly on data wrangling, EDA, statistical and machine learning based predictive models though I did include a Python based simulation of the Monte Hall 3-door problem. In teaching EDA using tools like ggplot2, I would challenge students to think about how they would go about trying to do some of the plots we were making in Excel. There is also some limited coverage of using Python to open Excel files and do some sheet manipulation. However, there was no real attempt to do the kind of modeling that was the focus of textbooks like PMS or DADM.

@@

Limitations of spreadsheets such as reproducibility and automation.

>>>>

Despite numerous calls for and predictions of the eradication of spreadsheets from the business world, Excel and other spreadsheets are alive and well. Yes, spreadsheets have well known limitations. Spreadsheets are not databases and probably should not be used as one, especially for relational data models. There are better data visualization and exploration tools. There are better dashboarding tools that are more easily maintained in a production environment. Spreadsheet errors are a huge problem. Version control and documentation present challenges. There are data size limitations and external data links present their own set of challenges. But, spreadsheets are still very well suited to ``what if?'' type modeling due to their flexibility, numerous built in functions, and their integrated sensitivity analysis, goal seeking and visualization capabilities. And of course, they are ubiquitous and familiar to hordes of analysts and modelers.

But, the analytics world has been changing. The advent of programmatic analytics using tools like R and Python have changed the practice of analytics as well as the teaching landscape \cite{bibid}. Limitations of spreadsheets with respect to reproducibility and automation have certainly played a role.  Additionally, both R and Python are free, open source, and have a rich set of analytics related libraries that have contributed to their growing popularity. Python, being a general purpose programming language, also offers advantages when moving models and analytical analysis pipelines from the prototype stage into production. Enormous data science ecosystems have emerged around both R and Python providing a wealth of learning resources for both new and experienced analytics students and professionals.

Numerous R and Python based data science courses have been developed and offered at universities and online learning platforms such as Coursera, EdX, Udemy and others. Some of these courses have made a large amount of their course content freely available for self-learners and for other educators. At our institution, we created a course called Practical Computing for Data Analytics (PCDA) in 2015 that introduced both R and Python (along with a little Linux) for use in data analytics work. The course is taught using combination of R Markdown documents \cite{bibid} and Jupyter notebooks \cite{bibid}. All of the course content is freely available from a publicly accessible course web site \url{URL}.

The PCDA course focuses mostly on data wrangling, exploratory data analysis (EDA), and statistical/machine learning based predictive models. It does include a Python based simulation of the Monte Hall 3-door problem. In teaching EDA using tools like ggplot2, I would challenge students to think about how they would go about trying to do some of the plots we were making in Excel. There is also some limited coverage of using Python to open Excel files and do some sheet manipulation. However, there was no real attempt to do the kind of modeling that was the focus of spreadsheet based modeling textbooks.

Literate computing \cite{Jupyter_proposal}. Educational narratives that support iterative exploration and development. The ability to weave standard language with formal programming languages is critical to creating the desired educational artifact. Jupyter used by many universities as a key component of their educational technology stack.

@@
